---
title: "CO~2~ MPIDR"
format: html
editor: source
reference-location: margin
---

```{r}
#| echo: false
library(data.table)
library(simplermarkdown)
library(ggplot2)
library(dplyr)
library(ggrepel)
#library(tidyverse)
library(forcats)

plot_data_dir <- file.path("..","..","Data")

rail_data_dir <- file.path("..","R_rail_distance")
load(file.path(rail_data_dir, "rail12geom.Rdata"))


# hint by nate, stackoverflow
auto_round_dt<- function(dt, ndigits=3, return_copy=TRUE){
  dt<- data.table::setDT(dt)
  roundme<- names(sapply(dt, class))[which(sapply(dt, class) == "numeric")]
  if(return_copy == TRUE){
    tmp<- data.table::copy(dt)
    out<- tmp[, (roundme):=round(.SD, ndigits), .SDcols=roundme]
    return(out)
  } else{
    return(dt[, (roundme):=round(.SD, ndigits), .SDcols=roundme])
  }
}
```

#### Rail distance 2019-2023

```{r}
#| output: asis
#| echo: false
rail <- auto_round_dt(rail12geom[,.(number=.N, distance=sum(distance)/.N, total=sum(distance)), by=.(purchase, year)], ndigits = 0, return_copy = TRUE)

cat(md_table(rail))

rail2 <- auto_round_dt(rail12geom[,.(number=.N, distance=sum(distance)/.N, total=sum(distance))], ndigits = 0, return_copy = TRUE)

rail2019 <- rail[purchase=="institute" & year == 2019,]
rail2020 <- rail[purchase=="institute" & year == 2020,]
rail2021 <- rail[purchase=="institute" & year == 2021,]
rail2022 <- rail[purchase=="institute" & year == 2022,]
rail2023 <- rail[purchase=="institute" & year == 2023,]

rail_mean_distance <- rail2$distance
```

purchase[^1], year, number (one-way), km Per Trip (distance)[^2], total distance in km (total)

[^1]: purchase definition\
    The institute booked most of the train tickets with the official travel partner (institute).\
    A small number of train tickets were booked by other means (offline).\
    Trips involving the use of monthly or annual tickets were not included. (BahnCard 100, Deutschlandticket).

[^2]: Distance Assumptions\
    The spherical distance was calculated. The actual distance is greater.

The mean rail distance (one-way) for 2019-2023 was `r rail_mean_distance` km.

## MPIDR 2019

### Data input

```{r}

CO2A <- data.table(year=rep(2019,9),cat=c(rep("Travel",7),rep("Energy",2)),
                    type=c("Flights: Westtours", "Flights: Self-booked (survey estimate)", "Train: Institute bookings", "Train: Self-booked (survey estimate)", "Bus: Self-booked (survey estimate)", "Car: Self-booked (survey estimate)", "Institute car", "Electricity", "Heating"),
                    rRate=c(100,16,100,16,16,16,100,100,100),
                    number=c(716,1,rail2019$number,21,5,2,1,NA,NA),
                    dist=c(2500,2500,rail2019$distance,rail_mean_distance,241,241,12199,NA,NA),
                    kWh=c(rep(NA,7),303000,449533),
                    factor=c(220,214,29,29,29,154,154,231,132.8))

CO2A[,CO2resp:=ifelse(!is.na(dist), number*dist*factor/1e6, kWh*factor/1e6)]
CO2A[,CO2all:=CO2resp * 100 / rRate]
```

Year, Category (cat), Type, Response Rate[^3] (rRate), Number (one-way), km Per Trip[^4] (dist), kWh, Factor[^5] (g CO~2~ / kWh or g CO~2~ / Pkm), t CO~2~ among respondents (CO2resp), t CO~2~ assuming 100% response rate (CO2all)

[^3]: Survey Response Rates\
    22 (61%) of the 36 survey participants had been at the institute in 2019.\
    89 (66%) of the 135 people working at the institute in 2019 were still at the institute in late 2021.\
    Our response rates were therefore 25% among the still-available people (22 / 89) and 16% among the whole 2019 population (22 / 135).

[^4]: Distance Assumptions\
    Average train distance (one-way, booked by institute): The spherical distance was calculated.\
    Average train distance (one-way, self-booked): The mean rail distance (one-way) for 2019-2023 was `r rail_mean_distance` km.\
    Average bus and car trip distance (one-way): Rostock - Berlin\
    Average flight distance (one-way): half of the 2019 average roundtrip distance provided by Westtours\
    For the institute car, the distance indicates the distance traveled over the entire year.

[^5]: Conversion Factors\
    Pkm = person kilometer; measures emissions per kilometer per individual passenger (rather than per vehicle)\
    Flight conversion factors are taken from Westtours' report, but not used here; instead we use the total tons of CO~2~ that they report.\
    Transport factors from German Environment Agency <https://www.umweltbundesamt.de/bild/vergleich-der-durchschnittlichen-emissionen-0>\
    Electricity conversion factors from "Energieträgermix übrige DREWAG-Produkte", Stadtwerke Dresden <https://drewag.de/wps/portal/drewag/cms/menu_main/privatkunden/produkte/strom/stromkennzeichnung>\
    Heating conversion factors from RSAG Rostock, "Fernwärmetarif ohne Klimakompensierung" <https://www.swrag.de/INTERSHOP/static/WFS/SWR-Portal-Site/-/SWR-Portal/de_DE/Veroeffentlichungen/Downloads/Fernwaerme/Zerti_FW309_CO2_Rostock_Zertifikat_2018.pdf>

### Results

```{r}
#| output: asis
#| echo: false
CO2A2<- auto_round_dt(dt=CO2A, ndigits = 2, return_copy = TRUE)
cat(md_table(CO2A2))
```

**The total CO~2~ footprint for 2021 was `r CO2A[,round(sum(CO2all),2)]` t.**

------------------------------------------------------------------------

## MPIDR 2020

### Data input

```{r}

CO2B <- data.table(year=rep(2020,9),cat=c(rep("Travel",7),rep("Energy",2)),
                    type=c("Flights: Westtours", "Flights: Self-booked (survey estimate)", "Train: Institute bookings", "Train: Self-booked (survey estimate)", "Bus: Self-booked (survey estimate)", "Car: Self-booked (survey estimate)", "Institute car", "Electricity", "Heating"),
                    rRate=c(100,19,100,19,19,19,100,100,100),
                    number=c(75,0,rail2020$number,8,0,0,1,NA,NA),
                    dist=c(4288,2500,rail2020$distance,rail_mean_distance,241,241,4756,NA,NA),
                    kWh=c(rep(NA,7),309886,388600),
                    factor=c(202.1,214,29,29,29,154,154,199,132.8))

CO2B[,CO2resp:=ifelse(!is.na(dist), number*dist*factor/1e6, kWh*factor/1e6)]
CO2B[,CO2all:=CO2resp * 100 / rRate]
```

Year, Category (cat), Type, Response Rate[^6] (rRate), Number (one-way), km Per Trip[^7] (dist), kWh, Factor[^8] (g CO~2~ / kWh or g CO~2~ / Pkm), t CO~2~ among respondents (CO2resp), t CO~2~ assuming 100% response rate (CO2all)

[^6]: Survey Response Rates\
    26 (72%) of the 36 survey participants had been at the institute in 2020.\
    89 (66%) of the 135 people working at the institute in 2020 were still at the institute in mid-2022.\
    Our response rates were therefore 29% among the still-available people (22 / 89) and 19% among the whole 2020 population (22 / 135).

[^7]: Distance Assumptions\
    See 2019.

[^8]: Conversion Factors\
    See 2019.

### Results

```{r}
#| output: asis
#| echo: false
CO2B2<- auto_round_dt(dt=CO2B, ndigits = 2, return_copy = TRUE)
cat(md_table(CO2B2))
```

**The total CO~2~ footprint for 2020 was `r CO2B[,round(sum(CO2all),2)]` t.**

------------------------------------------------------------------------

## MPIDR 2021

### Data input

```{r}

CO2C <- data.table(year=rep(2021,9),cat=c(rep("Travel",7),rep("Energy",2)),
                    type=c("Flights: Westtours", "Flights: Self-booked (survey estimate)", "Train: Institute bookings", "Train: Self-booked (survey estimate)", "Bus: Self-booked (survey estimate)", "Car: Self-booked (survey estimate)", "Institute car", "Electricity", "Heating"),
                    rRate=c(100,25,100,25,25,25,100,100,100),
                    number=c(33,2,rail2021$number,0,6,2,1,NA,NA),
                    dist=c(3290,2500,rail2021$distance,rail_mean_distance,241,241,1670,NA,NA),
                    kWh=c(rep(NA,7),309671,455000),
                    factor=c(220,214,29,29,29,154,154,199,132.8))

CO2C[,CO2resp:=ifelse(!is.na(dist), number*dist*factor/1e6, kWh*factor/1e6)]
CO2C[,CO2all:=CO2resp * 100 / rRate]
```

Year, Category (cat), Type, Response Rate[^9] (rRate), Number (one-way), km Per Trip[^10] (dist), kWh, Factor[^11] (g CO~2~ / kWh or g CO~2~ / Pkm), t CO~2~ among respondents (CO2resp), t CO~2~ assuming 100% response rate (CO2all)

[^9]: Survey Response Rates\
    33 (92%) of the 36 survey participants had been at the institute in 2021.\
    107 (82%) of the 131 people working at the institute in 2021 were still at the institute in mid-2022.\
    Our response rates were therefore 31% among the still-available people (33 / 107) and 25% among the whole 2021 population (33 / 131).

[^10]: Distance Assumptions\
    See 2019.

[^11]: Conversion Factors\
    See 2019.

### Results

```{r}
#| output: asis
#| echo: false
CO2C2<- auto_round_dt(dt=CO2C, ndigits = 2, return_copy = TRUE)
cat(md_table(CO2C2))
```

**The total CO~2~ footprint for 2021 was `r CO2C[,round(sum(CO2all),2)]` t.**

------------------------------------------------------------------------

## MPIDR 2022

### Data input

```{r}

CO2D <- data.table(year=rep(2022,5),cat=c(rep("Travel",3),rep("Energy",2)),
                    type=c("Flights: atmosfair", "Train: Institute bookings",
                           "Institute car", "Electricity", "Heating"),
                    rRate=c(100,100,100,100,100),
                    number=c(378,rail2022$number,1,NA,NA),
                    dist=c(2309,rail2022$distance,5732,NA,NA),
                    kWh=c(rep(NA,3),302806,394800),
                    factor=c(214,29,154,199,132.8))

CO2D[,CO2resp:=ifelse(!is.na(dist), number*dist*factor/1e6, kWh*factor/1e6)]
CO2D[,CO2all:=CO2resp * 100 / rRate]
```

Year, Category (cat), Type, Response Rate (rRate), Number (one-way), km Per Trip[^12] (dist), kWh, Factor[^13] (g CO~2~ / kWh or g CO~2~ / Pkm), t CO~2~ among respondents (CO2resp), t CO~2~ assuming 100% response rate (CO2all)

[^12]: Distance Assumptions\
    atmosfair reports 309 bookings for 2022 with a total of 872880 km of flights and a total of 186.3143 tonnes of C0~2~ (RFI 2.7). There where 309 bookings but probably at least 378 single flights (2309 km per single flight, 214g CO~2~ RFI 2.7 per person and km).

[^13]: Conversion Factors\
    See 2019.

### Results

```{r}
#| output: asis
#| echo: false
CO2D2<- auto_round_dt(dt=CO2D, ndigits = 2, return_copy = TRUE)
cat(md_table(CO2D2))
```

**The total CO~2~ footprint for 2022 was `r CO2D[,round(sum(CO2all),2)]` t.**

------------------------------------------------------------------------

## MPIDR 2023

### Data input

```{r}

CO2E <- data.table(year=rep(2023,5),cat=c(rep("Travel",3),rep("Energy",2)),
                    type=c("Flights: atmosfair", "Train: Institute bookings",
                           "Institute car", "Electricity", "Heating"),
                    rRate=c(100,100,100,100,100),
                    number=c(708,rail2023$number,1,NA,NA),
                    dist=c(2415,rail2023$distance,5612,NA,NA),
                    kWh=c(rep(NA,3),298321,407680),
                    factor=c(241.89,29,154,199,132.8))

CO2E[,CO2resp:=ifelse(!is.na(dist), number*dist*factor/1e6, kWh*factor/1e6)]
CO2E[,CO2all:=CO2resp * 100 / rRate]
```

Year, Category (cat), Type, Response Rate (rRate), Number (one-way), km Per Trip[^14] (dist), kWh, Factor[^15] (g CO~2~ / kWh or g CO~2~ / Pkm), t CO~2~ among respondents (CO2resp), t CO~2~ assuming 100% response rate (CO2all)

[^14]: Distance Assumptions\
    atmosfair reports 494 bookings for 2023 with a total of 1709587 km of flights and a total of 413.52 tonnes of C0~2~ (RFI 2.7). There where 494 bookings but probably 708 single flights (2415 km per single flight, 242g CO~2~ RFI 2.7 per person and km).

[^15]: Conversion Factors\
    See 2019.

### Results

```{r}
#| output: asis
#| echo: false
CO2E2<- auto_round_dt(dt=CO2E, ndigits = 2, return_copy = TRUE)
cat(md_table(CO2E2))
```

**The total CO~2~ footprint for 2023 was `r CO2E[,round(sum(CO2all),2)]` t.**

------------------------------------------------------------------------

```{r}
#| output: asis
#| echo: false

data_2019_2023 <- rbind(CO2A2,CO2B2,CO2C2,CO2D2,CO2E2)

#setwd("N:/Sustainability Group/internal/CO2/Data")

write.csv(data_2019_2023, file.path(plot_data_dir, "Data_2019_2023.csv"))

```

### Plots

```{r}
#| output: asis
#| echo: false

#setwd("N:/Sustainability Group/internal/CO2/Data")
dt <- read.csv(file.path(plot_data_dir, "Data_2019_2023.csv")) %>% 
  group_by(year) %>% 
  mutate(CO2tot=sum(CO2all))

# line plot with trends in CO2 by category

line_plot <- dt %>% 
  group_by(year,cat) %>% 
  summarise(CO2all=sum(CO2all),Total=CO2tot) %>% 
  distinct() %>% 
  ggplot()+
  geom_line(aes(year, CO2all,group_by=cat,col=cat),size=1) +
  geom_line(aes(year, Total,col="Total"), size=1)+
  scale_color_manual(values=c("black","#6699ff","#F46D43"),
                     breaks = c("Total", "Energy", "Travel"), name="Category")+
  theme_bw()+
  ylab("CO2 emissions (t)")+
  theme(plot.title = element_text(hjust = 0.5)) +
  ggtitle("CO2 emissions at the MPIDR by category, 2019-2023")
  

# pie chart of travel emissions

pie_plot <- dt %>% 
  group_by(cat) %>% 
  mutate(CO2tot=sum(CO2all)) %>%  
  filter(cat=="Travel") %>% 
  mutate(type=case_when(type=="Flights: Westtours"~"Flights",
                        type=="Flights: Self-booked (survey estimate)"~"Flights",
                        type=="Flights: atmosfair"~"Flights",
                        type=="Train: Institute bookings"~"Trains",
                        type=="Train: Self-booked (survey estimate)"~"Trains",
                        type=="Bus: Self-booked (survey estimate)"~"Car and Bus",
                        type=="Car: Self-booked (survey estimate)"~"Car and Bus",
                        type=="Institute car"~"Car and Bus",)) %>% 
  group_by(type) %>% 
  summarise(CO2all=sum(CO2all),CO2tot=CO2tot) %>% 
  distinct() %>% 
  mutate(prop=round((CO2all/CO2tot)*100,2),
         type=factor(type,levels=c("Trains","Car and Bus","Flights")),
         csum = rev(cumsum(rev(prop))), 
         pos = prop/2 + lead(csum, 1),
         pos = if_else(is.na(pos), prop/2, pos)) %>% 
   
  ggplot(aes(x="", y=prop, fill = fct_inorder(type))) +
  geom_col(width = 1, color = NA) +
  coord_polar(theta = "y",start=0) +
  scale_fill_manual(values=c("#FDAE61","#3288BD","#ABDDA4"), name="")+
  geom_label_repel(aes(y = pos, label = paste0(prop, "%")),
                   size = 4.5, nudge_x = 0, nudge_y=0, show.legend = FALSE) +
  guides(fill = guide_legend(title = "")) +
  theme_void()+
  theme(legend.text = element_text(size=12),plot.title = element_text(hjust = 0.5))+
  ggtitle("CO2 emissions at the MPIDR\nby means of transportation, 2019-2023")


#save plots 

ggsave(line_plot,file=file.path(plot_data_dir, "Plot_CO2tot.png"), width=2500,height = 1500,unit="px")
ggsave(pie_plot,file=file.path(plot_data_dir, "Plot_CO2bytransport.png"), width=1700,height = 1700,unit="px")

```
